Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-12-09 prop_letter_ajhs.tex

In this investigation, we explore the implications of extending
conventional quantum-mechanical probability measures to include the
effect of imprecise measurements.

There are long-standing debates in the foundations of quantum
mechanics regarding the tension between finite precision measurement
and contextuality~\cite{BarrettKent2004,Appleby_2005}. The outcome of
a theory that is contextual depends upon whether compatible sets of
observables are measured together or separately; a non-contextual
theory gives the same results in either case.  A classic example of
the varying opinions on the impact of imprecision is the disagreement
between Meyer~\cite{PhysRevLett.83.3751} and Mermin~\cite{Mermin1999}
on whether finite-precision measurements invalidate the spirit of the
Kochen-Specker
theorem~\cite{BELL_1966,kochenspecker1967,Redhead1987-REDINA,Mermin1990Simple,peres1995quantum,Jaeger2007,Held2016}.
The Kochen-Specker theorem is in essence a mathematical statement
about contextuality, asserting that in a Hilbert space of dimension
$d \ge 3$, it is impossible to associate determinate probabilities,
$\mu(P_i)= 0$ or $1$, with every projection operator~$P_i$, in such a
way that, if a set of commuting $P_i$ satisfies $\sum_i P_i = 1$, then
$\sum_i \mu(P_i) = 1$.  Meyer showed that one {\it can\/} assign
determinate probabilities $0$ or~$1$ to the measurement outcomes when
the Hilbert space is defined over the field of rational numbers,
therefore nullifying the theorem~\citep{PhysRevLett.83.3751}.
However, Mermin then argued that measurement outcomes must depend
smoothly on slight changes in the experimental configuration, leading
him to assert that the impact of Meyer's negative result is
``unsupportable'' on physical grounds~\citep{Mermin1999}.  The debate
on whether contextuality is consistent with experimental outcomes
continues to be an active topic of
research~\citep{Kent1999,HavlicekKrennSummhammerSvozil2001,SimonBruknerZeilinger2001,Cabello2002,Larsson2002,Appleby2002,BarrettKent2004,Appleby_2005,Spekkens2005,GuehneKleinmannCabelloEtAl2010,MazurekPuseyKunjwalEtAl2016}.

@@

%% ajh "begin" doubled, replace here; two missing commas
We start our investigation with the concept of $\delta$-determinism,
where~$\delta$ quantifies the effect of finite-precision uncertainties
on measurement outcomes. We recast the Kochen-Specker theorem using
$\delta$-determinism to quantify the effects of finite precision and
imperfections on contextuality. When $\delta=0$, the generalized
theorem reduces to the conventional one, but as $\delta$ varies from 0
to 1 we note a transition from contextuality to non-contextuality. For
a \emph{non-vanishing} range of $\delta$, quantum-mechanical
contextuality continues to hold, maintaining the Kochen-Specker
result, but at a \emph{certain fixed value}, $\delta = \frac{1}{3}$,
there is a sharp transition to \emph{non-contextuality}, parallel in
character to a phase transition.  These results provide a new insight
into the Meyer-Mermin debate by presenting a theory in which there is
a \emph{parameter} interpolating between what appeared previously to
be irreconcilable aspects of contextuality.

The quantum-mechanical requirement for contextuality
continues to hold, thus maintaining the Kochen-Specker result for a
\emph{non-vanishing} range of $\delta$, but that at a \emph{certain
  fixed value} of $\delta$, there is a sharp transition, parallel in
its consequences to, say, a phase transition, to
\emph{non-contextuality}. %% ; that is, suddenly a quantum system can be
%% simulated by a non-contextual hidden variable theory.
%% , turning the
%% system effectively from a quantum system into a classical system. 
% that is, quantum mechanics suddenly fails to hold, and all variables
% can be specified simultaneously, turning the theory effectively into a
% classical theory that can be specified consistently by hidden
% variables.
% \yutsung{The part after ``that is''
%   looks strange.
% \begin{enumerate}
% \item The Kochen-Specker theorem is not about quantum mechanics, is about whether
% quantum mechanics can be simulated by a non-contextual deterministic hidden
% variable theory or not. Hence, ``quantum mechanics suddenly fails to hold''
% doesn't make a lot sense.
% \item The question is whether there exist a QIVPM means ``a classical theory
% consistently by hidden variables.'' At least, a QIVPM is not deterministic, and
% I need to check whether it qualities as a classical probabilistic hidden variable
% theory\ldots(?)
% \end{enumerate}
% }
These results provide a new insight into the Meyer-Mermin debate by
presenting a theory in which there is a \emph{parameter} interpolating
between what appeared previously to be irreconcilable aspects %interpretations
of contextuality.
% ajh- above: I think "interpretations" is not quite right -> "aspects"?

@@

The quantum probability postulates assume a mathematical idealization
in which quantum states and measurements are both infinitely
precise. In an actual experimental setup with an ensemble of quantum
states that would ideally be identical, but are not actually
identically prepared, with imperfections and inaccuracies in measuring
devices, an experimenter might only be able to determine that the
probability of an event $P$ is concentrated in the range $[0.49,0.51]$
instead of being precisely $0.5$. The spread in this range depends on
the amount of resources (time, energy, money, etc.)  that are devoted
to the experiment. In the classical setting, this ``fuzziness'' can be
formalized by moving to \emph{interval-valued probability measures}
(IVPMs), which we explore in the next section, along with our proposed
extension to the quantum domain.

>>>>

In this investigation, we explore the implications of extending
conventional quantum-mechanical probability measures to include the
effect of imperfect measurements. 
There are long-standing debates in the foundations of quantum
mechanics regarding the tension between finite precision measurement
and contextuality~\cite{BarrettKent2004,Appleby_2005}. The outcome of
a theory that is contextual depends upon whether compatible sets of
observables are measured together or separately; a non-contextual
theory gives the same results in either case.  A classic example of
the varying opinions on the impact of imprecision is the disagreement
between Meyer~\cite{PhysRevLett.83.3751} and Mermin~\cite{Mermin1999}
on whether finite-precision measurements invalidate the spirit of the
Kochen-Specker
theorem~\cite{BELL_1966,kochenspecker1967,Redhead1987-REDINA,Mermin1990Simple,peres1995quantum,Jaeger2007,Held2016}.
The Kochen-Specker theorem is in essence a mathematical statement
about contextuality, asserting that in a Hilbert space of dimension
$d \ge 3$, it is impossible to associate determinate probabilities,
$\mu(P_i)= 0$ or $1$, with every projection operator~$P_i$, in such a
way that, if a set of commuting $P_i$ satisfies $\sum_i P_i = 1$, then
$\sum_i \mu(P_i) = 1$.  Meyer showed that one {\it can\/} assign
determinate probabilities $0$ or~$1$ to the measurement outcomes when
the Hilbert space is defined over the field of rational numbers,
therefore nullifying the theorem~\citep{PhysRevLett.83.3751}.
However, Mermin then argued that measurement outcomes must depend
smoothly on slight changes in the experimental configuration, leading
him to assert that the impact of Meyer's negative result is
``unsupportable'' on physical grounds~\citep{Mermin1999}.  The debate
on whether contextuality is consistent with experimental outcomes
continues to be an active topic of
research~\citep{Kent1999,HavlicekKrennSummhammerSvozil2001,SimonBruknerZeilinger2001,Cabello2002,Larsson2002,Appleby2002,BarrettKent2004,Appleby_2005,Spekkens2005,GuehneKleinmannCabelloEtAl2010,MazurekPuseyKunjwalEtAl2016}.

@@

%% ajh "begin" doubled, replace here; two missing commas
We start our investigation with the concept of $\delta$-determinism,
where~$\delta$ quantifies the effect of finite-precision uncertainties
on measurement outcomes. We recast the Kochen-Specker theorem using
$\delta$-determinism to quantify the effects of finite precision and
imperfections on contextuality. When $\delta=0$, the generalized
theorem reduces to the conventional one, but as $\delta$ varies from 0
to 1 we note a transition from contextuality to non-contextuality. For
a \emph{non-vanishing} range of $\delta$, quantum-mechanical
contextuality continues to hold, maintaining the Kochen-Specker
result, but at a \emph{certain fixed value}, $\delta = \frac{1}{3}$,
there is a sharp transition to \emph{non-contextuality}, parallel in
spirit to a phase transition.  These results provide a new insight
into the Meyer-Mermin debate by presenting a theory in which there is
a \emph{parameter} interpolating between what appeared previously to
be irreconcilable aspects of contextuality.

%The quantum-mechanical requirement for contextuality
%continues to hold, thus maintaining the Kochen-Specker result for a
%\emph{non-vanishing} range of $\delta$, but that at a \emph{certain
%  fixed value} of $\delta$, there is a sharp transition, parallel in
%its consequences to, say, a phase transition, to
%\emph{non-contextuality}. %% ; that is, suddenly a quantum system can be
%% simulated by a non-contextual hidden variable theory.
%% , turning the
%% system effectively from a quantum system into a classical system. 
% that is, quantum mechanics suddenly fails to hold, and all variables
% can be specified simultaneously, turning the theory effectively into a
% classical theory that can be specified consistently by hidden
% variables.
% \yutsung{The part after ``that is''
%   looks strange.
% \begin{enumerate}
% \item The Kochen-Specker theorem is not about quantum mechanics, is about whether
% quantum mechanics can be simulated by a non-contextual deterministic hidden
% variable theory or not. Hence, ``quantum mechanics suddenly fails to hold''
% doesn't make a lot sense.
% \item The question is whether there exist a QIVPM means ``a classical theory
% consistently by hidden variables.'' At least, a QIVPM is not deterministic, and
% I need to check whether it qualities as a classical probabilistic hidden variable
% theory\ldots(?)
% \end{enumerate}
% }
%These results provide a new insight into the Meyer-Mermin debate by
%presenting a theory in which there is a \emph{parameter} interpolating
%between what appeared previously to be irreconcilable aspects %interpretations
%of contextuality.
% ajh- above: I think "interpretations" is not quite right -> "aspects"?

@@

The quantum probability postulates assume a mathematical idealization
in which quantum states and measurements are both infinitely
precise, i.e., {\it sharp}. 
In an actual experimental setup with an ensemble of quantum
states that would ideally be identical, but are not actually
identically prepared, with imperfections and inaccuracies in measuring
devices, an experimenter might only be able to determine that the
probability of an event $P$ is concentrated in the range $[0.49,0.51]$
instead of being precisely $0.5$. The spread in this range depends on
the amount of resources (time, energy, money, etc.)  that are devoted
to the experiment. In the classical setting, this ``fuzziness'' can be
formalized by moving to \emph{interval-valued probability measures}
(IVPMs), which we explore in the next section, along with our proposed
extension to the quantum domain.
